---
title: "Process Visualisation"
author: "Nima Ramezani"
# date: "27/07/2021"
output: html_document
---

```{r setup, include=FALSE, echo = F}
knitr::opts_chunk$set(warning = F)
library(magrittr)
library(dplyr)

library(bupaR)
library(eventdataR)

library(rbig)
library(rutils)
library(rvis)
library(rprom)

```

# Introduction

This article gives you an overview of some visualisations that you can generate with *rprom*.

We are using 
package [eventdataR](https://cran.r-project.org/web/packages/eventdataR/index.html)
to get some sample data for visualisation.
We may use various datasets from this package.

First of all, we build an empty *Transition System* object:

```{r build_ts, eval=T}
tsobj = new('TransitionSystem')
```

Now, we pick an eventlog from `eventdataR` package and feed our object with the data:

```{r get_data, eval=T}

traffic_fines_df = eventdataR::traffic_fines %>% as.data.frame()

traffic_fines_df %>% head %>% knitr::kable()
```

We can define completion of an activity as status for our cases, 
so we filter for `lifecycle` to be `complete`.
You also need to specify critical column headers for the data feeder:

```{r feed_obj, eval=T}
tsobj$feed.transition_events(traffic_fines_df %>% filter(lifecycle == 'complete'), 
                    caseID_col = 'case_id', 
                    status_col = 'activity',
                    startTime_col = 'timestamp')
```

# Process maps:

The first and most valuable process visualisation is a *process map*. 

Process map is a directed graph containing nodes and edges. 
Each node is associated with a *status* and each edge denotes a *transition* from one status to another.

Method `plot.process.map` can be used to plot the process map for the transition system:

```{r pmap_rate, eval=T}
tsobj$plot.process.map()
```

## Process Map Measures:

### Rate

Each edge (link or transition) in the process map has a *source status* and a *destination status*.
By default method `plot.process.map` uses *rate* as the measure denoted by edges. 
On each edge, 
you will see the the rate (ratio) of transitions from the source status going to the destination status 
as a percentage of total outgoing transitions from the source status.
The number denoted by each node(status) shows the total incoming transition to that status.

### Frequency 

When you choose frequency as measure, the incoming edges of each status annotate frequencies of *entries* to that status while the outgoing edges annotate the *exit* frequencies (outgoing from that status to other statuses). 
This option can be chosen by setting argument `measure` to `freq`.

### Time

Other than frequencies in the process map, you can also use *time* as measure to focus on the durations of statuses and transitions. You can specify the time unit via argument `time_unit`:

```{r pmap_freq, echo=T}
tsobj$plot.process.map(measure = 'time', time_unit = 'hour')
```

Values denoted on each status(node) show average duration of lingering in that status and values denoted on the edges show average transition times which is average duration of lingering in the source status before transiting to the destination status.

## Case Filtering:

Sometimes showing all transitions in the process map, makes it look noisy and complicated.
We may need to filter out some cases with unusual traces in order to have a better
visualisation of the process.

The most common filtering is to filter out cases which have uncommon transitions.
These are what we call *noisy cases*. 
These cases follow traces different from the most frequent traces of the process and thus deviate from the main process.


Case filtering can be done via method `filter.case`. 
To filter out noisy cases, you can set argument `FreqRateCut` to a value between 0 and 1.
Usually, a value close to 1.0 is chosen. For example, if you set it to 0.9, 
10% of cases with the least frequent traces will be eliminated:

```{r pmap_filter_frc}
tsobj$set.filter.case(freq_rate_cut = 0.9)

tsobj$plot.process.map()
```

When you case filter your object, all measures and charts will be impacted by the filtering until 
you run method `filer.reset` or run `filter.case` with different filtering arguments.
For complete information about case filtering refer to the package API reference.
<!-- todo:  -->
<!-- - Add a link to the API reference. -->
<!-- - Complete documentation for method filter.case -->

## Process Map Plotters:

*rprom* currently supports two plotters for rendering process map graph: `grviz` and `visNetwork`. 
The default plotter is `grviz` which is using javascript package [GraphVis](http://www.graphviz.org) and is available via the R package [rviz](https://github.com/genpack/rvis).
To use plotter `visNetwork`, you need to have R package [visNetwork](https://datastorm-open.github.io/visNetwork/) installed:

```{r install_visNetwork, eval=F}
if(!require(visNetwork)){install.packages('visNetwork')}
```

`visNetwork` plots become noisy and unreadable when the graph size is big. 
So it's best to use a frequancy rate cut filtering before plotting to filter out cases with low-frequent traces and simplify the process:

```{r pmap_visNetwork}
tsobj$reset.filters()
tsobj$set.filter.case(freq_rate_cut = 0.8)

tsobj$plot.process.map(plotter = 'visNetwork')
```

The layout of the plot generated by `grvis` looks better, on the other side, 
`visNetwork` gives you possibility of dragging the nodes or enable physics.

You can customize your plot by specifying parameters to the `config` argument.

For example, for the `visNetwork` plot, 
you can select `hierarchical` layout, change direction, edge smoothing type and enable physics for nodes and edges:

```{r pmap_visNetwork_customized}

tsobj$reset.plots()
customized = list(layout = 'hierarchical', node.physics.enabled = T, direction = 'left.right', 
                  link.smooth = list(enabled = T, type = 'curvedCCW'))

tsobj$plot.process.map(plotter = 'visNetwork', config = customized,  width = "800px", height = "1200px")
```


You can refer to the documentation of `rvis` package to see how you can specify argument `config` in order to customize your plot. 
`rvis` uses a unique configuration format for all the plotters it supports. 
You can also directly use `visNetwork` functions to customize your plot. 
For example you can change edge arrow sizes.

```{r pmap_visNetwork_customized2, eval=F}
tsobj$reset.plots()

tsobj$plot.process.map(plotter = 'visNetwork') %>% 
  visNetwork::visEdges(arrows = list(to = list(enabled = TRUE, scaleFactor = 1.5)))
```

For complete instructions refer to the documentation of `visNetwork` package.


# Process Sankey

Sankey chart is another type of visualisation which can show you the process by 
focusing more on the process flows.

```{r plot.process_sankey}
tsobj %>% plot_process_sankey()
```


# Process Tree

Another type of visualisation ideal for showing a process is an interactive Sankey-Tree.

To generate this visualisation, you will need to install packages `sankeytreeR` and `treemap`.
package `treemap` can be installed from cran and `sankeytree` needs to be installed from github:

```{r install_sankey_tree}
if(!require(treemap)){install.packages('treemap')}
if(!require(d3r)){install.packages('d3r')}
if(!require(sankeytreeR)){devtools::install_github('https://github.com/timelyportfolio/sankeytree.git')}

tsobj %>% plot_process_tree()
```




