---
title: "Random Walk on Process Graph"
author: "Nima Ramezani"
date: "16/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = F)
library(magrittr)
library(dplyr)
library(lubridate)

library(rbig)
library(rutils)
library(rvis)
library(rprom)
```

## Introduction

A random walk on a graph is a process that begins at some status (vertex, node or state), and 
at each time step moves to another status. 

When the graph edges are unweighted, 
the transition destination status is chosen uniformly at random among the neighbors of the current status. 
cases transition to a neighbor status with probability proportional to the weight of the corresponding edge. 
While the case traces (list of statuses in the order they are visited) of a particular random walk is sometimes of interest, 
it is often more productive to reason about the expected behavior of a random walk. 
To this end, we will investigate the probability distribution over the statuses after a certain number of steps.

In this article, we use *rprom* to study random walks on a process map represented by a directed graph. 

## Transition Matrix

We will let the vector $\vec{p}^{t} \in \mathbb{R}^n$ denote the probability vector at time $t$ and
$p^{(t)}_{A}$ denote the probability of being in state $A$ at time $t$.  
The status probabilities must add up to $1.0$.

To obtain the status probability distribution after each transition, we simply mutiply the **Transition Matrix** also known as *Adjacency Matrix* by the current probability vector: $$\vec{p}^{(t)} * A = \vec{p}^{(t+1)}$$

<!-- <img src="https://render.githubusercontent.com/render/math?math=e^{i \pi} = -1"> -->

### Get Transition Matrix from Transition System objects

Here, we will show you how to obtain te Transition(Adjacency) Matrix from an onject of class `TransitionSystem`.

We used dataset `hospital_billing` from package eventdata for this study 
as it contains case status data in column `state`.
To shorten the status labels, we renamed them to their abbreviated form.
You can make any changes to the eventlog history of a `TransitionSystem` object directly but you will need to 
reset the object after that: 

```{r feed_data}

hospital_df = eventdataR::hospital_billing %>% as.data.frame
status_mapper = c('In progress' = 'IP', 'Closed' = 'CL', 'Released' = 'RE', 'Billed' = 'B', 
                  'Invoice rejected' = 'IR', 'Billable' = 'BL', 'Empty' = 'E', 'Rejected' = 'RJ',
                  'Unbillable' = 'UB', 'Check' = 'C')
mapped_rows = which(hospital_df$state %in% intersect(hospital_df$state, names(status_mapper)))
hospital_df$state[mapped_rows] = status_mapper[hospital_df$state[mapped_rows]]

tsobj = new('TransitionSystem')
tsobj$feed.transition_events(hospital_df, 
                    caseID_col = 'case_id', 
                    status_col = 'state',
                    startTime_col = 'timestamp')
```

Let's have a look at the process map for this Transition System:

```{r pmap_rate, eval=T}
tsobj$plot.process.map(width = "1000px", height = "800px")
```

The value of cell in row $i$ and column $j$ of matrix A
is the probability of the walk at status $i$ selecting
the edge to status $j$.

You can get the **Transition Matrix** of the process represented by a Transition System  
by calling method `get.transition_matrix`.

```{r adjacency}
 tsobj$get.transition_matrix()
```

## Random Walk Steps

So let's consider one case which is in status **IP (In Progress)** and 
see how is its status likely to change in the next steps.

The initial probability vector is $\vec{p}^{(0)} = (0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)$ in which 
the second element associated with status **IP** is $1$.
The probability distribution after one step is:

```{r step_1}
A  = tsobj$get.transition_matrix()

p0 = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
p1 = p0 %*% A
print(p1)
```

which is the second row of the Transition Matrix associated with status **IP**.  
As we take further steps, the probability distribution changes. 
For example to see the probability distribution after 10 steps,
we multiply the transition matrix 10 times recursively:

```{r step_10}
p = list(p1)
for (i in 1:9){
  p[[i + 1]] = p[[i]] %*% A
}
print(p[[10]])
```

`TransitionSystem` object has a method which computes the random walk probability distributions given the initial state and number of steps:

```{r random_walk, results = 'asis'}
knitr::kable(tsobj$random_walk(starting_status = 'IP', num_steps = 10))
```


Looking at the probabilities of each status versus the step number, 
we can see which statuses cases are pulled towards status **END** as the step number increases.

```{r probs_versus_month}
tsobj$random_walk(starting_status = 'IP', num_steps = 10) %>% 
  rvisPlot(x = list(step_no = 1:10), y = as.list(rownames(A)), plotter = 'plotly', type = 'scatter', shape = 'line.point')

```


## Periodic Transition System

In the previous study, we could see how probability distribution among statuses change as we walk through the process graph. 
We could see which statuses are most likely met in each stage of the process, however
this does not give us any view of how the process status changes over time because
status transitions can happen at any time and durations cases spend at statuses can be very different.

**Periodic Transition Systems**, help us consider time in the modelling of the process.
If the time interval between transitions are constant, number of steps represents time elapsed and
we will be able to monitor status probabilities over time.

Periodic Transition System assumes a status transition at the start of a period of time like a day for example.
As you know we can always have a transition from one status to itself, so at the beginning of each time period (day),
a transition happens: if the case is still in the same status as yesterday, a same-status-transition event is triggered.

The TransitionSystem object has a method named `to_periodic` which converts itself into a periodic transition system and returns it as a new object.
We can use that object to generate the transition matrix and observe probability distributions over time.
Let's see how this works:


```{r probs_versus_time}

ts_monthly = tsobj$to_periodic('month')

ts_monthly$random_walk(starting_status = 'IP', num_steps = 10) %>% 
  rvisPlot(x = list(month_no = 1:10), y = as.list(ts_monthly$get.statuses()), plotter = 'plotly', type = 'scatter', shape = 'line')


```


Well, you can see how the probabiolity distribution is being pulled towards the END status as time increases.

In the next study, we will learn how to run a simulation to monitor the future of the process.


